\chapter{Ethics and Data Management}
\label{edm}
\section{Ethical Considerations}
There are some ethical considerations needed. 
Phages can be used to treat bacterial infections and may need to be administered under the supervision of a doctor. 
Attempting to control phages in food or the environment by releasing a phage cocktail into waterways could cause issues further down the line if the released solution contains unwanted chemicals. 
An imbalance can be introduced into the ecosystem, further exacerbating the problem. 
An additional step in the food production process will increase food costs and make food production more difficult to control. 
The cost of creating, maintaining, and using phages at an industrial scale can become substantial and require a significant amount of energy. 
Dumping phages into the ecosystem could cause issues if the phage concoction includes resources that the bacteria can use, and this can become costly for taxpayers. 

\section{Data Management}
All data and results can be found on \href{https://github.com/BiggusVickus/Master-Thesis}{GitHub}. 
Some simulation data will have to be recreated as the \textit{.parquet} data files are too big for GitHub to store. 
Measures have been taken to label and document the code, datasets, and parameter configurations, for example, as shown in \Cref{AppendixE}. 
Any qualified researcher should be able to replicate, audit, use, and edit the computational experiments and code if needed. 
This systematic approach to version control and storage aligns with best practices, ensuring that results are both traceable and verifiable. 

\section{Adherence to Codes and Principles}
I acknowledge that the thesis adheres to the \href{https://student.uva.nl/en/topics/ethics-in-research}{ethical code} and \href{https://rdm.uva.nl/en}{research data management policies} of UvA and IvI.

The following table lists the data used in this thesis, with the source code. 
I confirm that the list is complete and that the listed data are sufficient to reproduce the results of the thesis. 

\begin{table}[ht!]
    \begin{tabular}{|l|l|l|}
    \hline
    \textbf{\begin{tabular}[c]{@{}l@{}}Short description \\ (max. 10 words)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Availability \\ (e.g., URL, DOI)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}License \\ \end{tabular}} \\ \hline
        Dataset & \href{https://github.com/BiggusVickus/Master-Thesis/tree/main/SimulationResults}{Simulation Results} & MIT \\ \hline 
        Source Code & \href{https://github.com/BiggusVickus/Master-Thesis}{Source Code} & MIT \\ \hline
        Simulation Conditions & \Cref{AppendixE} and text under figures and in the text& \\ \hline
    \end{tabular}
\end{table} 